Unreliable heterogeneity: how measurement error obscures heterogeneity in meta-analyses in psychology

### 2023-06-18 manuscript/manuscript.qmd
osf links to figures/tables

```{r Figure 1, out.width= "85%"}
#| fig.env: figure*
#| fig-pos: 't'
#| fig-cap: "Figure 1. Flowchart summarizing the simulation design."

@@

```{r Figure 2, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/manuscript/z-r-hs-plot.pdf")
```

_Figure 2. Unreliability in primary studies leads to a net negative bias when heterogeneity is positive. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson’s r or translation to Fisher’s z these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line corresponds to an average reliability in primary studies. Results are for standard deviation in reliability = 0.15, within-study sample size = 150, and 20 studies within each meta-analysis. Code to reproduce figure: LINK._

@@

```{r Figure 3, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/manuscript/r_tau_0.02-0.08.pdf")
```

_Figure 3. Unreliability in primary studies typically leads to a net negative bias even for small degrees of heterogeneity. Effect size is Pearson’s r and rows indicate the number of studies in the meta-analysis. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to effect size truncation in Pearson’s r these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line corresponds to an average reliability in primary studies. Results are for standard deviation in reliability = 0.15, within-study sample size = 150. Code to reproduce figure: LINK._

@@

```{r Figure 4, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/manuscript/sample_size.pdf")
```

_Figure 4. A smaller sample size in primary studies tends to exacerbate the bias in observed heterogeneity estimates due to unreliability. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications. Rows indicate the (fixed) sample size within studies and columns indicate nominal true between-studies standard deviation. Due to effect size truncation in Pearson's $r$ these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line corresponds to an average reliability in primary studies. Results are for standard deviation in reliability = 0.15 and 20 studies within each meta-analysis._

>>>>

```{r Figure 1, out.width= "90%"}
#| fig.env: figure*
#| fig-pos: 't'
#| fig-cap: "Flowchart summarizing the simulation design. For product-moment correlations, option (1) corresponds to the Hedges and Vevea (1998) model and option (2) to the Schmidt and Hunter (2015) model (see methods section 'Meta-analytic model'). $\\theta$ = average true effect size (i.e., without sampling or measurement error), $\\tau^2$ = variance of true effect sizes, $\\theta_i$ = study true effect size, $\\gamma_i$ = study effect size with sampling error $\\sigma_i^2$, $N_i$ = study sample size, $R_i$ = study reliability, $\\bar{R}$ = average reliability across studies, $\\delta$ = standard deviation of reliabilities across studies, $r_i$ = observed study effect size with sampling and measurement error, $\\bar{r}$ = average observed effect size across studies. This figure was created using the website draw.io."

@@

```{r Figure 2, echo=FALSE, out.width = '90%'}

#| fig-cap: "Unreliability in primary studies leads to a net negative bias when heterogeneity is positive. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson’s r or translation to Fisher’s z these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line corresponds to an average reliability in primary studies. Results are for standard deviation in reliability = 0.15, within-study sample size = 150, and 20 studies within each meta-analysis. Code to reproduce figure: osf.io/8ygwj."

knitr::include_graphics("../figures/manuscript/z-r-hs-plot.pdf")
```

@@

```{r Figure 3, echo=FALSE, out.width = '90%'}
#| fig-cap: "Unreliability in primary studies typically leads to a net negative bias even for small degrees of heterogeneity. Effect size is Pearson’s r and rows indicate the number of studies in the meta-analysis. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to effect size truncation in Pearson’s r these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line corresponds to an average reliability in primary studies. Results are for standard deviation in reliability = 0.15, within-study sample size = 150. Code to reproduce figure: osf.io/8ygwj."

knitr::include_graphics("../figures/manuscript/r_tau_0.02-0.08.pdf")
```

@@

```{r Figure 4, echo=FALSE, out.width = '90%'}
#| fig-cap: "A smaller sample size in primary studies tends to exacerbate the bias in observed heterogeneity estimates due to unreliability. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications. Rows indicate the (fixed) sample size within studies and columns indicate nominal true between-studies standard deviation. Due to effect size truncation in Pearson's $r$ these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line corresponds to an average reliability in primary studies. Results are for standard deviation in reliability = 0.15 and 20 studies within each meta-analysis. Code to reproduce figure: osf.io/8ygwj."

knitr::include_graphics("../figures/manuscript/sample_size.pdf")
```
